---
title: "R Coding Exercise"
output:
  html_document:
    toc: FALSE
---

```{r}
#load packages and check data
library("dslabs")
library("tidyverse")
library("ggplot2")
#look at help file for gapminder data
help(gapminder)
#get an overview of data structure
str(gapminder)
#get a summary of data
summary(gapminder)
#determine the type of object gapminder is
class(gapminder)
```

#Write code that assigns only the African countries to a new object/variable called africadata

```{r}
africadata <- gapminder %>% filter (continent== "Africa")
```

#overview of africadata

```{r}
str(africadata)
```

#summary of africadata

```{r}
summary(africadata)
```

#first new object: infant mortality and life expectancy data for African countries

```{r}
africa_il <- 
africadata %>% 
select(infant_mortality,life_expectancy)
```

#second new object: life expectancy and population data for African countries

```{r}
africa_lp <- 
africadata %>%
select(life_expectancy,population)
```

#overview of africa_il and africa_lp

```{r}
str(africa_il)
str(africa_lp)
```

#summary of africa_il and africa_lp

```{r}
summary(africa_il)
summary(africa_lp)
```

#plot life expectancy as a function of infant mortality

```{r}
africa_il %>%
ggplot() + aes(x= infant_mortality,y= life_expectancy) + geom_point()
```

#plot life expectancy as a function of population size

```{r}
africa_lp %>%
ggplot() + aes(x=log10(population), y=life_expectancy) + geom_point() 
```

#identify NA values in infant_mortality

```{r}
subset(africadata,is.na(infant_mortality)) 
```

#since there is NA data prior to 1981 and after 2016, choose a year like 2000 to avoid any NAs #make a dataset only including data from 2000

```{r}
year_2000 <-africadata %>% filter(year == "2000")
```

#get overview and summary of year_2000

```{r}
str(year_2000)
summary(year_2000)
```

#new object: infant mortality and life expectancy data for African countries IN 2000

```{r}
africa_il_2000 <- 
year_2000 %>% 
select(infant_mortality,life_expectancy)
```

#new object: life expectancy and population data for African countries IN 2000

```{r}
africa_lp_2000 <- 
year_2000 %>%
select(life_expectancy,population)
```

#plot life expectancy as function of infant mortality in 2000 only

```{r}
africa_il_2000 %>% 
ggplot() + aes(x= infant_mortality,y= life_expectancy) + geom_point()
```

#plot life expectancy as a function of population size with 2000 data

```{r}
africa_lp_2000 %>%
ggplot() + aes(x=log10(population), y=life_expectancy) + geom_point() 
```

#fit 2000 data to a linear model (outcome = life_expectancy, predictor = infant_mortality)

```{r}
il.fit <- lm(life_expectancy ~ infant_mortality, data=year_2000)
summary(il.fit) 
```

#statistically significant correlation (p=2.83e-8)

#fit 2000 data to a linear model (outcome = life_expectancy, predictor = population)

```{r}
lp.fit <- lm(life_expectancy ~ population, data=year_2000)
summary(lp.fit) 
```

#no statistically significant correlation (p=0.61)


## **this section added by Yao Lu**

```{r}
library('DataExplorer')
library(ggplot2)
```
**data clean**
```{r}
a <- gapminder

plot_missing(a)
```

There 28% missing for gdp, try to narrow the continent to asia to see if that get better
```{r}
asiadata <- a[which(a$continent=='Asia'),] 
plot_missing(asiadata)
```  

Not better. Here we have 3 choice. Do imputation manually, like substitute missing GDP per capital by the mean of GDP per capital.
Use 'MICE' package to generate a new data set with automatic imputation.
Drop the rows with missing gdp.

Here for simplicity, I choose drop the missing.

```{r}
asiadata1 <- asiadata[which(!is.na(asiadata$gdp)),]
plot_missing(asiadata1)
asiadata2 <- asiadata1[which(!is.na(asiadata1$infant_mortality)),]

asiadata2$gdppercap <- asiadata2$gdp/asiadata2$population
```

**data visualization**

```{r}

table(droplevels(asiadata2$region))

hist(asiadata2$infant_mortality)
hist(asiadata2$life_expectancy)
hist(asiadata2$fertility)
hist(asiadata2$gdppercap)
```

```{r}
p <- ggplot(data = asiadata2, aes(x = year, y = infant_mortality, group= country, color=country))
p + geom_line()+ggtitle("infant_mortality") +  xlab("year") + ylab("life time")

```
Here we can see all the infant_mortality decrease as the time goes on with different decreasing rate.

```{r}
p <- ggplot(data = asiadata2, aes(x = year, y = life_expectancy, group= country, color=country))
p + geom_line()+ggtitle("infant_mortality") +  xlab("year") + ylab("life time")

```
Here we can see all the life_expectancy increase as the time goes on with different increasing rate.

**data modeling**

```{r}
lm1 <-lm(asiadata2$life_expectancy~asiadata2$infant_mortality+asiadata2$fertility+asiadata2$region+asiadata2$gdppercap)
anova(lm1)
```
Here we can see that all the variables are significant. For detailed relation direction, we will see below.


```{r}
summary(lm1)
```
Here we know infant mortality and fertility have negative influence on life time. Gdp per capital have positive influence on life time. 

The life time in different regions are not all same.

```{r}
#save the output
outcome <- broom::tidy(lm1)
```

**future work we can do**

Do imputation to make use of the entire data. Find the countries who have the top life time to find what's their similarities.