---
<<<<<<< HEAD
title: "Module 4 Data Analysis Exercise"
output: 
  html_document:
    toc: FALSE
---
=======
title: "Week 4 Data Analysis Exercise"
author: "Irene Cavros"
editor: visual
---


>>>>>>> main
***Summary***

Data for this exercise was pulled from the CDC's data repository. [This dataset](https://data.cdc.gov/NCHS/NCHS-Drug-Poisoning-Mortality-by-County-United-Sta/pbkm-d27e) describes drug poisoning deaths at the county level by selected demographic characteristics and includes age-adjusted death rates for drug poisoning from 1999 to 2015.

**Loading packages, dataset, and reviewing the raw data:**

<<<<<<< HEAD
=======
---

>>>>>>> main
```{r}
#load packages
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(naniar)

#load dataset
<<<<<<< HEAD
drug_poison_data <- read_csv("dataanalysis-exercise/data/raw/drug_poison_data.csv")
=======
drug_poison_data <- read_csv("dataanalysis-exercise/data/raw_data/drug_poison_data.csv")
>>>>>>> main

#examine data
summary(drug_poison_data)

```

<<<<<<< HEAD
=======
---
>>>>>>> main
**Creating a new dataframe called "data" with variables of interest:**

The selected variables here are year, state, county, population and death rate.

```{r}
#refined dataset with only variables of interest
data <- drug_poison_data %>%
    select(year, state, county, population, death_rate)
```

**Filter to look at only cases in Georgia:**

```{r}
georgia <- data %>%
  filter(state %in% "Georgia")
<<<<<<< HEAD
=======

>>>>>>> main
```

This new final dataset (georgia) gives us cases of drug poisoning from 1999-2015 in the state of Georgia.

**Save cleaned data as an RDS file in processed_data folder:**

```{r}
<<<<<<< HEAD
saveRDS(georgia, file="dataanalysis-exercise/data/processed/georgia_clean.rds")
=======
saveRDS(georgia, file="dataanalysis-exercise/data/processed_data/georgia_clean.rds")
>>>>>>> main
```

**Save summary table:**

```{r}
#make summary table
summary_georgia = data.frame(do.call(cbind, lapply(georgia, summary)))
print(summary_georgia)

#save it as an RDS file
saveRDS(summary_georgia, file="dataanalysis-exercise/results/summary_georgia.rds")
<<<<<<< HEAD

```

=======
```

---
>>>>>>> main
