---
title: "Module 8 Exploration"
output: html_document
---

This Quarto file loads the cleaned data and does some exploring.

As part of the exploratory analysis, this code produces plots, tables, and other summary quantities for some interesting aspects of our data

#### Load packages

```{r}
library(tidyverse) #for transforming data
library(here) #to set paths for data loading/saving
library(dplyr) #for data processing/cleaning
library(tidyr) #for data processing/cleaning
library(gtsummary) #for summarizing data
library(forcats) #for working with categorical variables
library(ggplot2) #for nice plots
library(cowplot) #addon to ggplot for nice plots
library(skimr) #for nice visualization of data 
library(tidymodels) #for modeling

```

####Load the data

```{r}
#Path to data. Note the use of the here() package and not absolute paths
cleandata <- readRDS(here("fluanalysis", "data", "cleandata.rds"))
cleandata
```
#### Look at variable types
```{r}
str(cleandata)
```

#### Produce and save a summary table.

This also helps us get things into the shape of a data frame (for easier saving)

```{r}
# looking at the data
summary_df <- dplyr::glimpse(cleandata)
# save to file
summarytable_file = here("fluanalysis", "data", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

#### Outcomes of interest

For our further analysis, we decide that our main continuous outcome of interest is **Body temperature**, and our main categorical outcome is **Nausea**. We want to see if the other symptoms are correlated with (predict) those outcomes.

```{r}
summary(cleandata)
```

Let's produce and print some summary output with a smaller subset of variables. 

```{r}
# make subset with a few variables to summarize
subset <- cleandata %>% select(BodyTemp, SwollenLymphNodes, ChillsSweats, AbPain, Diarrhea, Nausea, Vomit)

# summarize the data
subsettable <- 
  tbl_summary(subset)
subsettable
```

#### Outcome of interest #1: Body temperature

Since body temperature is continuous, let's create a histogram/density plot.

```{r}
p1 <- ggplot(subset, aes(BodyTemp)) + 
  geom_density() + 
  labs(title = "Body Temperature Density", x = "Body Temperature (Farenheit)", y = "Density") + scale_fill_discrete(labels=c('No', 'Yes')) + theme_classic()

p1
```
This tells us the frequency of body temperatures in the dataset. 

Let's create some boxplots including this outcome and the shortlist of independent variables/predictors we have chosen.

```{r}
#boxplot of BodyTemp x SwollenLymphNodes
p2 <- ggplot(subset, aes(x=BodyTemp, y=SwollenLymphNodes)) + 
  geom_boxplot(fill = "gray") +
  labs(title = "BodyTemp x SwollenLymphNodes", x = "Body Temperature (Farenheit)", y = "Swollen Lymph Nodes") + scale_fill_discrete(labels=c('No', 'Yes')) + theme_classic()

p2
```
Based on this plot, it looks like the average body temperature among those without swollen lymph nodes looks marginally higher with a larger spread. This doesn't really point to anything too interesting so let's try chills. 

```{r}
#boxplot of BodyTemp x ChillsSweats
p3 <- ggplot(subset, aes(x=BodyTemp, y=ChillsSweats)) + 
  geom_boxplot(fill = "gray") +
  labs(title = "BodyTemp x ChillsSweats", x = "Body Temperature (Farenheit)", y = "Experienced Chills or Sweats") + scale_fill_discrete(labels=c('No', 'Yes')) + theme_classic()

p3
```
In this plot, the mean body temperature among those who experienced chills or sweating appears about a half degree higher than those without, and the spread is larger as well. This could potentially be interesting, but let's do a few more plots to look at the nausea outcome as well. 

For now, let's use nausea as a predictor and keep bodytemp as the outcome. 

```{r}
#boxplot of BodyTemp x Nausea
p4 <- ggplot(subset, aes(x=BodyTemp, y=Nausea)) + 
  geom_boxplot(fill = "gray") +
  labs(title = "BodyTemp x Nausea", x = "Body Temperature (Farenheit)", y = "Nausea") + scale_fill_discrete(labels=c('No', 'Yes')) + theme_classic()

p4
```
Nothing appears too interesting here. 

####  Outcome of interest #2: Nausea

Next let's make nausea the outcome of interest. We can create some plots for this outcome and a few independent variables/predictors.

```{r}
#plot of Nausea x AbPain
NauAb <- data.frame(table(subset$Nausea,subset$AbPain))
names(NauAb) <- c("Nausea","AbPain","Count")

p5<- ggplot(data=NauAb, aes(x=Nausea, y=Count, fill=AbPain)) + geom_bar(stat="identity")
p5
```
Abdominal pain looks to be more common in those with nausea, than those without, which makes sense!

```{r}
#plot of Nausea x Diarrhea
NauDi <- data.frame(table(subset$Nausea,subset$Diarrhea))
names(NauDi) <- c("Nausea","Diarrhea","Count")

p6<- ggplot(data=NauDi, aes(x=Nausea, y=Count, fill=Diarrhea)) + geom_bar(stat="identity")
p6
```
Diarrhea also appears to be more frequent among those with nausea, which, again, is logical. 

```{r}
#plot of Nausea x Vomit
NauVom <- data.frame(table(subset$Nausea,subset$Vomit))
names(NauVom) <- c("Nausea","Vomit","Count")

p7<- ggplot(data=NauVom, aes(x=Vomit, y=Count, fill=Nausea)) + geom_bar(stat="identity")
p7
```
This time we changed the X axis so we are now looking at among those who vomited, how many had nausea and among those who did not vomit, how many had nausea? A large proportion of those who vomited also had nausea, which also makes sense. 






