---
title: "Module 8 Wrangling"
output: 
  html_document:
    toc: FALSE
---

#### Introduction

During the exercise for model 8, we will practice a bit of model fitting using the `tidymodels` framework.

#### Load packages

```{r}
library('tidyverse') #for transforming data
library('dplyr') #for data processing/cleaning
library('tidyr') #for data processing/cleaning
library('skimr') #for nice visualization of data 
library('here') #to set paths
```

#### Get the data

Note: the file we want to work with is actually an RDS file/object, not an RData/RDA file, despite the somewhat misleading .Rda file name ending. So you need to use the right function to load RDS files into R.

```{r}
data <- readRDS(here("fluanalysis", "data", "SympAct_Any_Pos.Rda"))
data
```

#### Remove all variables that have Score or Total or FluA or FluB or Dxname or Activity in their name.

Don't do this manually one by one, figure out how to use R commands that let you remove things in an efficient manner, using e.g., `contains()` or `starts_with()` from `dplyr`/`tidyselect`

Also remove the variable `Unique.Visit.`

```{r}
data <- as.data.frame(data)
df2 <- data %>% 
    select(-contains('Score')) %>% #to remove variables containing xyz
    select(-contains('Total')) %>%
    select(-contains('FluA')) %>%
    select(-contains('FluB')) %>%
    select(-contains('Dxname')) %>%
    select(-contains('Activity')) %>%
    select(-c('Unique.Visit')) #to remove final variable 
```

We are now left with 32 variables coding for presence or absence of some symptom. Only one, temperature, is continuous. A few have multiple categories.

#### Let's remove any NA observations

There aren't many.

```{r}
cleandata<- na.omit(df2)
cleandata
```

#### Saving the cleaned data

```{r}
save_data_location <- here::here("fluanalysis", "data", "cleandata.rds")
saveRDS(cleandata, file = save_data_location)
```

